Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-08-28 Manuscript/manuscript.Rmd
wip on methods, unfinished

The complete dataset used for modeling was composed of two components: the response vector **Y**, an indicator variable separating between TAD and non-TAD boundary regions, and the predictor-type specific data matrix, **X** = {$X_1$, $X_2$, ..., $X_p$}, where $p$ was the complete set of predictors. The data was split into a 7:3 ratio of training and testing sets.

### DATA RE-SAMPLING

Class imbalance was (CI) defined as the ratio of minority class observations to the sum of minority and majority class observations:

@@

where $C_{min}$ denotes the set of TAD boundary regions and $C_{maj}$ denotes the complementary set non-TAD boundary regions. Here, CI is bounded by 0 and 0.50. Heavily imbalanced classes tend to approach 0, while balanced classes approach 0.50.

In order to assess the influence of imbalanced classes, we evaluated four re-sampling techniques (including to no re-sampling). They are described in detail below.

1. **No Sampling**: All of the data points from the majority and minority classes of the training set were used.

2. **Random Under-Sampling (RUS)**: All of the data points from the minority class of the training set were used. Sampling without replacement was used to obtain the same number of data points from the majority class.

3. **Random Over-Sampling (ROS)**: All of the data points from the majority class of the training set were used. Sampling with replacement was used to obtain the same number of data points from the minority class.

@@

### FINAL MODEL ASSESSMENT AND PERFORMANCE

>>>>

The complete dataset used for modeling was composed of two components: the response vector **Y**, an indicator variable separating between TAD and non-TAD boundary regions, and the predictor-type specific data matrix, **X** = {$X_1$, $X_2$, ..., $X_p$}, where $p$ was the set of predictors. One, or all types of predictors were used as described. The data was split into a 7:3 ratio of training and testing sets.

### Addressing class imbalance

Class imbalance (CI) was defined as the proportion of minority class out of the total number of observations:

@@

where $C_{min}$ denotes the set of TAD boundary regions and $C_{maj}$ denotes the set non-TAD boundary regions. CI of heavily imbalanced classes approaches 0, while CI for balanced classes approaches 0.50. To assess the effect of CI on predictive performance, we first used the data with **o Sampling**, that is, the majority and minority classes were used as is. We evaluated three re-sampling techniques: **Random Under-Sampling (RUS)** - ll of the data points from the minority class of the training set were used. Sampling without replacement was used to obtain the same number of data points from the majority class.

3. **Random Over-Sampling (ROS)**: All of the data points from the majority class of the training set were used. Sampling with replacement was used to obtain the same number of data points from the minority class. [??? I leave this unfinished because the definitions seem plain wrong]

@@

[??? Stopped here]

### Assessmend of model performance
